---
title: "Assign"
author: "Samuel Mweni"
date: "2/15/2023"
output: word_document
---



```{r}
library(tidyverse)
library(ggplot2)
library(dplyr)
```
#1 Use and show r code to import the Astronauts csv file.
```{r}

astron<- read.csv("Astronauts.csv")
head(astron)

```

#2 Use and show r code to determine the number of Astronauts who are
# in the data table and the number of descriptive column variables.

```{r}
dim(astron)
```
There are 357 astronauts and 19 descriptive column variables.


# 3 Use and show r code to produce the number and the
# names of only the female Astronauts.  
```{r}
astron%>%
  filter(Gender=="Female")->fastron
fastron
```
# 4 Use and show R coding to produce a table that shows the
# military rank and the military branch of each Female Astronaut.
```{r}
fastron%>%
  select(Military.Rank,Military.Branch)
```
# 5 Some of the Astronauts in the data table are listed as Active 
# and others are not. Use and show r code to determine how many 
# Female Astronauts were currently active when this data set was
# constructed
```{r}
fastron%>%
  filter(Status=="Active")->Hfastron
count(Hfastron)
```
# 6 Use and show ggplot coding to produce a bar graph that compares
# space walks for Male and Female Astronauts
```{r}
astron%>%
  select(Gender,Space.Walks)->Mastron
```

```{r}
# Create the bar graph
ggplot(Mastron, aes(x = Gender, y = Space.Walks, fill = Gender)) +
  geom_bar(stat = "identity", width = 0.5) +
  ggtitle("Number of Spacewalks by Male and Female Astronauts") +
  ylab("Number of Spacewalks") +
  xlab("Gender")


```
# 7 Use and show r code to produce the table below that lists all
# Female Astronauts who have registered one or more Space Walks.
```{r}
fastron%>%
  filter(`Space.Walks`>= 1)%>%
  group_by(Name,Gender, `Space.Walks`)%>%
  arrange(Name)%>%
  summarise(n=n())
  
```


# 8  Use and show ggplot coding to produce a box plot in order to 
# investigate outliers and the overall distribution for Space
# Walks for males and female astronauts combined. Does the boxplot 
# indicate that there are outliers? Using the boxplot, estimate the
# IQR for the data.
```{r}
ggplot(astron, aes(x = "", y = Space.Walks)) +
  geom_boxplot() +
  ggtitle("Box Plot of Spacewalks  by Gender") +
  ylab("SpaceWalks") +
  xlab("Gender")

IQR(astron$Space.Walks)
```
Yes, there are existence of outliers

# 9 Use and show ggplot coding to produce side by side box plots
# in order to investigate outliers and the overall distributions
# for Space Walks for males and female astronauts separately
```{r}
ggplot(astron, aes(x = Gender, y = Space.Walks)) +
  geom_boxplot() +
  ggtitle("Box Plot of Spacewalks  by Gender") +
  ylab("SpaceWalks") +
  xlab("Gender")
```



# 10 The histogram displayed below shows that the Space Walks data
# is generally skewed right for all Astronauts (males and females).
# Use and show R code that will produce the histogram displayed.
```{r}
ggplot(astron, aes(x = Space.Walks)) +
  geom_histogram(binwidth = 1.2, color = "black", fill = "red") +
  ggtitle("The histogram displayed below shows that the Space Walks ") +
  xlab("Space walks") +
  ylab("Frequency")

```


# 11 Use and show R coding to calculate the average(mean) of
# Space Walks for Male and Female Astronauts separately.
```{r}

mean(fastron$Space.Walks)
```
```{r}
astron%>%
  filter(Gender=="Male")->MaleAstron
mean(MaleAstron$Space.Walks)
```
# 12 Does a linear relationship exist between Space Flights and
# Space Walks ? Investigate by using and showing ggplot coding to
# produce a scatter plot.  Comment on your findings based on the
# configuration of the scatter plot data points.
```{r}
ggplot(astron, aes(x = Space.Walks, y =Space.Flights)) +
  geom_point() +
  ggtitle("Scatter Plot of X vs Y") +
  xlab("Space walsks") +
  ylab("Space flight")

```
There is no linear relationship between spacewalks and space flights 

# 13 Mae C. Jemison was the first African American female Astronaut.
# Use and show R code that will output her Undergraduate Major,
# Graduate Major, and Alma Mater.
```{r}
fastron%>%
  filter(Name=="Mae C. Jemison")%>%
  select(Undergraduate.Major,Graduate.Major,Alma.Mater)
  

```
# 14 Use and show R code to determine how many female
# Astronauts were members of the US Air Force or retired
# from the US Air Force at the time that the data set was
# generated.
```{r}
astron2 <- fastron%>%
  filter(Gender=="Female")%>%
  filter(`Military.Branch`=="US Air Force" | `Military.Branch`=="US Air Force (Retired)")
count(astron2)

```


